Clustering time series based on probability distributions across temporal granularities

### 2021-10-29 main.Rmd
characterisation of clusters subsection done

Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis. A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings. @Cook2007-qe provides several ways to characterize clusters.

(a) Parallel coordinate plot: Parallel coordinate plot (@wegman1990hyperdimensional) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns. In a Parallel Coordinates Plot, each variable has its own axis, which are all parallel. Each axis is connected by a series of lines. That is, each line is made up of connected points on each axis. The order of the axes can affect how the reader interprets the data. This is because adjacent variables are more easily perceived than non-adjacent variables and rearranging the axes can reveal patterns or correlations between variables. Scattered plots of the p variables are arranged in a scatterplot matrix. It's a neat way to show multiple relationships at once, and it allows us to compare all the plots at once.

(b) Scatterplot matrix: The scatter plot matrix (draftsman's plot) is a matrix that comprises pairwise scatter plots of the p variables.Pairwise scatter plots are excellent for determining relationships between variables and determining which factors have contributed the most to clustering.

(c) Plotting cluster statistics: For larger problems, parallel coordinate plots may become cluttered and difficult to read, therefore we may opt to display cluster statistics instead. (@Dasu_undated-gw)

(c) MDS, PCA and t-SNE: While all of the techniques examine a matrix of distances or dissimilarities to give a representation of the data points in a reduced-dimension space, their goals are not the same. The principal component analysis (Johnson & Wichern 2002) attempts to retain data variance. Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space. t-SNE, on the other hand, is concerned with preserving neighbourhood data points. Close data points in high-dimensional space will be condensed in the t-SNE embeddings.

(d) Tour: A tour (@wickham2011tourr) is a collection of interpolated linear projections of multivariate data into a lower-dimensional space. The sequence is seen as a dynamic visualization, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
<!-- and also provides examples on how one might spin and brush using a tour, to build up a cluster analysis manually. -->

Depending on the distance measure utilized for the study, the cluster characterization technique will differ. Clusters that utilize wpd-based distances are characterised using multi-dimensional scaling and parallel coordinate displays. For JS-based distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met. This technique may potentially make advantage of multi-dimensional scaling.

>>>>

\noindent Cluster characterization, both quantitatively and qualitatively, is a crucial aspect in cluster analysis. @Cook2007-qe lists numerous methods for characterising clusters. Listed below are a few techniques and R packages that are utilized in this study.

(a) _Parallel coordinate plots_ (@wegman1990hyperdimensional) are often used to visualise high-dimensional and multivariate data, allowing visual grouping and pattern detection.. A Parallel Coordinates Plot features parallel axes for each variable. Each axis is linked by lines. The axes' arrangement may affect the reader's interpretation of the data. Changing the axes may reveal patterns or relationships between variables for categorical variables. However, for categories with cyclic temporal granularities, preserving the underlying ordering is more desirable.

(b) _Scatterplot matrix_ contains pairwise scatter plots of the p variables. Pairwise scatter plots are useful for figuring out how variables relate to each other and how factors determine the clustering.

(c) _Displaying cluster statistics_ are useful when we have larger problems and it is difficult to read the Parallel coordinate plots due to congestion. (@dasu2005grouping)

(d) _MDS, PCA and t-SNE_ While all of them use a distance or dissimilarity matrix to construct a reduced-dimension space representation, their goals are diverse. PCA seeks to retain data variance (Johnson and Wichern 2002). Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space. t-SNE, on the other hand, is concerned with preserving neighbourhood data points. The t-SNE embeddings will compress data points which are close in high-dimensional space.

(e) _Tour_ is a collection of interpolated linear projections of multivariate data into lower-dimensional space. As a result, the viewer may observe the high-dimensional data's shadows from a low-dimensional perspective.

The cluster characterization approach varies depending on the distance metric used. Parallel coordinate plots, scatter plot matrices, MDS or PCA are potentially useful ways to characterize clusters using wpd-based distances. For JS-based distances, plotting cluster statistics is beneficial for characterization and variable importance could be displayed through parallel coordinate plots. This part of the work uses R packages `GGally` (@R-GGally), `Rtsne` (@R-tsne), `ggplot2` (Wickham2009pk), `tour` (@wickham2011tourr), `stats` (@R-language).
